---
title: "DATA2002"
format: 
  html: 
    self-contained: true
    code-fold: true
    ### USEFUL ###
    code-tools: true # Includes a menu to show/hide all chunks
    ### OPTIONAL ###
    code-line-numbers: true # Line numbers in code chunks
    df-print: paged # Sets how dataframes are automatically printed
    theme: cerulean # Controls the font, colours, etc.
table-of-contents: true # (Useful) Creates a table of contents!
number-sections: true # (Optional) Puts numbers next to heading/subheadings
---

# Data Tidying

```{r,message=FALSE,warning=FALSE}
library(tidyverse)
library(dplyr)

data <- read.delim("bodyfat.txt")

# arrange data in ascending order 
data_order <- data %>% 
 arrange(Pct.BF) 

# override weight and height columns in pounds and inches to kilograms and meters, categorise ages, create BMI variable
data1 <- data_order %>% 
  mutate(Height = (Height/39.37)) %>%
  mutate(Weight = (Weight/2.205)) %>%
  mutate(BMI = (Weight/(Height)^2))

# only keep percentage body fat values greater than 3% (remove two points)
df <- subset(data1, Pct.BF>=3)

# remove density
df <- df[ , !names(df) %in% c("Density")] ## works as expected

```

## Pairs Plot

```{r}
suppressWarnings({ 
  panel.cor <- function(x, y){
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- round(cor(x, y), digits=2)
    txt <- paste0("R = ", r)
    text(0.5, 0.5, txt)}

upper.panel <- function(x, y){
  points(x,y)}


pairs(df[,c(1:4, 7:9, 16)], 
      lower.panel = panel.cor,
      upper.panel = upper.panel)
})
```

# Models

Unless the number of candidate variables \> sample size (or number of events), use a backward stepwise approach (<https://quantifyinghealth.com/stepwise-selection/>)

Josh mentioned his group used weight and BMI (indirectly includes height)?? - neither model output includes either variable

## Backward stepwise selection

```{r}
# intercept-only model
intercept_only <- lm(Pct.BF ~ 1, data = df)

# model with all predictors
model <- lm(Pct.BF ~ ., data = df) #Added intercept back in

# backward stepwise regression
backward <- step(model, direction = 'backward', scope = formula(model), trace = TRUE)

# results 
backward$anova

# final model
backward$coefficients

summary(backward) #Added summary to examine P values
```

Model:

pct.bf \~

## Backward stepwise selection (plot method/same result)

```{r,warning=FALSE,message=FALSE}
library(olsrr)

model <- lm(Pct.BF ~ ., data = df)

ols_step_backward_aic(model)
a <- ols_step_backward_aic(model)

# stepwise backward regression plot
plot(a)

# final model 
a$model

```

Model:

pct.bf \~ 449.55 - 415.14\*density + 0.01500\*age + 0.04816\*chest + 0.04140\*hip - 0.05793\*bicep

## Forward stepwise selection

```{r}

# intercept-only model
intercept_only <- lm(Pct.BF ~ 1, data = df)

# model with all predictors
model <- lm(Pct.BF ~ ., data = df)

# forward stepwise regression
forward <- step(intercept_only, direction = 'forward', scope = formula(model), trace = TRUE)

forward <- step(intercept_only, direction = 'forward', scope = formula(model), trace = 0)

# results 
forward$anova

# final model
forward$coefficients

```

Model:

pct.bf \~ 445.22 - 409.17\*density + 0.057464\*abdomen + 0.011260\*age

```{r}
# if cannot implement, install.package('ggfortify')
library(ggfortify)
# determine log(k) k Value
nrow(df)
M1 <- lm(Pct.BF ~ ., data = df) 

# Doing BIC
M1.step <- step(M1, 
                scope = list(lower = df ~ 1, 
                             upper = df ~ .),
               direction = "backward", k = log(248))

# sumary of bic
summary(M1.step)
# graph of bic
autoplot(M1.step)
```

# Performance

```{r}
# Forward AIC model
# PCT.bf ~  Waist + Weight + Wrist + BMI + Age + Chest
forward_value = round(summary(forward)$r.squared,3) 

# Backward AIC model
# PCT.bf ~ Age + Neck + Chest + Abdomen + Hip + Forearm + Wrist + BMI
backward_value = round(summary(backward)$r.squared,3)
```

## In sample

The forward AIC model has an R-squared value of `r forward_value` while the backward model has an value of `r backward_value`. It seems that backward AIC models fits the dataset a little better.

```{r}
sjPlot::tab_model(forward,backward,show.ci = FALSE,dv.labels = c("Forward model", "Backward model"))
```

## Out of sample

We use 10 fold cross validation here over 5 fold since there are 248 observations in the dataset and we allow more observations to train the model. Using 10 fold cross validation, the backward AIC model has slight advantage over forward AIC as shown below.

```{r,message=FALSE}
library(caret)
set.seed(1)

# 10 fold cross validation
ten_fold = trainControl(
    method = "cv", number = 10,
    verboseIter = FALSE
  )

# forward model
cv_forward = train(
  Pct.BF ~  Waist + Weight + Wrist + BMI + Age + Chest, df,
  method = "lm",
  trControl = ten_fold
)

cv_backward = train(
  Pct.BF ~ Age + Neck + Chest + Abdomen + Hip + Forearm + Wrist + BMI, df,
  method = "lm",
  trControl = ten_fold
)
Method = c("forward AIC","backward AIC")
RMSE = c(cv_forward$results$RMSE,cv_backward$results$RMSE)
MAE  = c(cv_forward$results$MAE,cv_backward$results$MAE)
knitr::kable(data.frame(Method, RMSE,MAE))


```

```{r}
set.seed(1)
results <- resamples(list(Forward_AIC = cv_forward, Backward_AIC = cv_backward))

# put into a dataframe
mae = c(results$values$`Forward_AIC~MAE`,results$values$`Backward_AIC~MAE`)
rmse = c(results$values$`Forward_AIC~RMSE`,results$values$`Backward_AIC~RMSE`)
combine = data.frame(mae,rmse)
combine = combine%>% mutate(type = c(rep("Forward_AIC",10),rep("Backward_AIC",10)))
colnames(combine) <- c('MAE','RMSE','type')

# draw the plot
p1 = plotly::ggplotly(ggplot(combine,aes(x = type,y=MAE,color=type)) +geom_boxplot()+
  labs(y = "MAE")+theme_light()+theme(axis.title.y = element_blank())+coord_flip())

p2 = plotly::ggplotly(ggplot(combine,aes(x = type,y=RMSE,color=type)) +geom_boxplot()+
  labs(y = "RMSE")+theme_light()+theme(axis.title.y = element_blank())+coord_flip())

manipulateWidget::combineWidgets(p1, p2, nrow = 2)
```

```{r}
# # 10 fold cross validation
# ten_fold = trainControl(
#     method = "cv", number = 5,
#     verboseIter = FALSE
#   )
# 
# # forward model
# cv_forward_five = train(
#   Pct.BF ~  Waist + Weight + Wrist + BMI + Age + Chest, df,
#   method = "lm",
#   trControl = ten_fold
# )
# 
# cv_backward_five = train(
#   Pct.BF ~ Age + Neck + Chest + Abdomen + Hip + Forearm + Wrist + BMI, df,
#   method = "lm",
#   trControl = ten_fold
# )
# Method = c("forward AIC","backward AIC")
# RMSE = c(cv_forward_five$results$RMSE,cv_backward_five$results$RMSE)
# MAE  = c(cv_forward_five$results$MAE,cv_backward_five$results$MAE)
# knitr::kable(data.frame(Method, RMSE,MAE))

```











