---
title: "Merge data"
output: html_document
date: "2023-05-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load data & packages

```{r Loading in packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(janitor)
library(sf)
```
# merge data for developing model
## read in fish data
```{r Loading in the data}
catch_95 <- read.csv("Data/CatchInd1995_1999.csv") %>% clean_names()
catch_00 <- read.csv("Data/CatchInd2000_2004.csv") %>% clean_names()
catch_05 <- read.csv("Data/CatchInd2005_2009.csv") %>% clean_names()
catch_10 <- read.csv("Data/CatchInd2010_2014.csv") %>% clean_names()

cell <- read_xlsx("Data/Codes.xlsx", sheet = "Cell") %>% clean_names()
gear <- read_xlsx("Data/Codes.xlsx", sheet = "Gear") %>% clean_names()
taxa <- read_xlsx("Data/Codes.xlsx", sheet = "Taxa") %>% clean_names()
country <- read_xlsx("Data/Codes.xlsx", sheet = "Country") %>% clean_names()

```

## read in index data
```{r}
index <- read.csv("Data/IndexInd.csv") %>% clean_names()
```

## filtering Indonesian fish records, and combining across the different year ranges.
```{r Filtering for Indonesian records}
indo_code <- country %>% filter(fao_name == "Indonesia") %>% pull(country)
tuna_code <- taxa %>% filter(common_name == "Yellowfin tuna") %>% pull(taxon_key) # sample to select Yellowfin tuna in indonesia
indo_tuna_ids <- index %>% filter(c_number == indo_code, taxonkey == tuna_code) %>% pull(id)

catch_95_filtered <- catch_95 %>% filter(id %in% indo_tuna_ids)
catch_00_filtered <- catch_00 %>% filter(id %in% indo_tuna_ids)
catch_05_filtered <- catch_05 %>% filter(id %in% indo_tuna_ids)
catch_10_filtered <- catch_10 %>% filter(id %in% indo_tuna_ids)

catch_all <- rbind(catch_95_filtered, catch_00_filtered, catch_05_filtered, catch_10_filtered)
```

## Merge fish data from different year
1. Merge in the location, year, and gear data.
2. Filter down to the Indonesian region (we were only considering Indonesian vessels before).
3. Aggregate results across fishing gears.
```{r Merging the data}
catch_cleaned <- catch_all %>% 
  left_join(cell, by = "cell") %>% 
  left_join(index[, c("id", "i_year", "gear")], by = "id") %>% 
  filter(lat_centre >= -10, lat_centre <= 13, lon_centre >= 93, lon_centre <= 134) %>% 
  group_by(lat_centre, lon_centre, i_year) %>% 
  summarise(sum_reported = sum(reported),
            area = mean(ocean_areasqkm)) %>% 
  ungroup() %>% 
  arrange(lat_centre, lon_centre, i_year) %>% 
  rename(latitude = lat_centre, longitude = lon_centre, year = i_year) %>% 
  mutate(rate_reported = round(1000*sum_reported/area, 4)) %>% # Landings per 1000 square kilometres
  select(-sum_reported, -area)
```

## load in the reef data from the labs and filter it to have the same range as the fishing data.
```{r Loading in the reef data}
reef <- read.csv("Data/Reef_Check_with_cortad_variables_with_annual_rate_of_SST_change.csv") %>% 
  clean_names() %>% 
  filter(latitude_degrees >= min(catch_cleaned$latitude), latitude_degrees <= max(catch_cleaned$latitude), 
         longitude_degrees >= min(catch_cleaned$longitude), longitude_degrees <= max(catch_cleaned$longitude),
         year >= min(catch_cleaned$year), year <= max(catch_cleaned$year)) %>% 
  select(reef_id, latitude_degrees, longitude_degrees, year, average_bleaching, clim_sst) %>% 
  rename(latitude = latitude_degrees, longitude = longitude_degrees)
```

## Perform a nearest-neighbour join between the reef and catch datasets.
```{r Nearest-neighbour join}

merged <- data.frame()

for (curr_year in unique(reef$year)) {
  catch_sf <- catch_cleaned %>% 
    filter(year == curr_year) %>% 
    as.data.frame() %>% 
    st_as_sf(coords = c("longitude", "latitude"))
  
  reef_sf <- reef %>% 
    filter(year == curr_year) %>% 
    st_as_sf(coords = c("longitude", "latitude"))

  # Compute distances between each point in catch_sf and reef_sf
  distances <- st_distance(catch_sf, reef_sf)

  # Get the minimum distance for each catch point
  min_distances <- apply(distances, 1, min)

  # Join the datasets based on nearest features
  merged_year <- st_join(catch_sf, reef_sf, join = st_nearest_feature) %>% 
    as.data.frame() %>% 
    mutate(latitude = st_coordinates(.$geometry)[, 2], longitude = st_coordinates(.$geometry)[, 1]) %>% 
    select(-year.x, -year.y, -geometry) %>% 
    mutate(year = curr_year)

  # Add the minimum distance as a new column to the merged_year dataframe
  merged_year$distance_to_nearest_reef <- min_distances

  rownames(merged_year) <- NULL
  merged <- rbind(merged, merged_year)
}
```

## Load in and merge the SAU fishing effort data, before normalising the catch data by fishing effort.
```{r Fishing effort data}
effort <- read.csv("Data/SAU Effort FishingEntity 83 v50-1.csv") %>% clean_names() %>% 
  filter(fishing_sector == "Industrial") %>% 
  group_by(year) %>% 
  summarise(sum_effort = sum(effort)) %>% 
  mutate(sum_effort = sum_effort/1e+06)

merged_effort <- merged %>% 
  left_join(effort, by = "year") %>% 
  filter(!is.na(sum_effort)) %>% 
  mutate(rate_norm = rate_reported/sum_effort) %>% 
  select(-rate_reported)

write.csv(merged_effort, file = "merged.csv")
```

# Merge data for iterative map
```{r}
merged <- read.csv("merged.csv")
reef_data <- read.csv("Data/Reef_Check_with_cortad_variables_with_annual_rate_of_SST_change.csv") %>%
  clean_names() %>%
  select(reef_id, latitude_degrees, longitude_degrees)


merged_with_reef_coords <- left_join(merged, reef_data, by = "reef_id")


unique_reefs <- merged_with_reef_coords %>%
  distinct(reef_id, latitude_degrees, longitude_degrees)


filtered_merged_with_reef_coords <- merged_with_reef_coords %>%
  group_by(reef_id, year) %>%
  slice_min(distance_to_nearest_reef, n = 1) %>%
  ungroup()

reef_merged <- data.frame()

for (curr_year in unique(reef$year)) {
  catch_sf <- catch_cleaned %>% 
    filter(year == curr_year) %>% 
    as.data.frame() %>% 
    st_as_sf(coords = c("longitude", "latitude"))
  
  reef_sf <- reef %>% 
    filter(year == curr_year) %>% 
    st_as_sf(coords = c("longitude", "latitude"))

  # Compute distances between each point in catch_sf and reef_sf
  distances <- st_distance(reef_sf, catch_sf)

  # Get the minimum distance for each reef point
  min_distances <- apply(distances, 1, min)

  # Join the datasets based on nearest features
  reef_merged_year <- st_join(reef_sf, catch_sf, join = st_nearest_feature) %>% 
    as.data.frame() %>% 
    mutate(latitude = st_coordinates(.$geometry)[, 2], longitude = st_coordinates(.$geometry)[, 1]) %>% 
    select(-year.x, -year.y, -geometry) %>% 
    mutate(year = curr_year)

  # Add the minimum distance as a new column to the reef_merged_year dataframe
  reef_merged_year$distance_to_nearest_catch <- min_distances

  rownames(reef_merged_year) <- NULL
  reef_merged <- rbind(reef_merged, reef_merged_year)
}

merged_effort <- reef_merged %>% 
  left_join(effort, by = "year") %>% 
  filter(!is.na(sum_effort)) %>% 
  mutate(rate_norm = rate_reported/sum_effort) %>% 
  select(-rate_reported)

write.csv(merged_effort, file = "reef_merged.csv")
```

# Merge data for important fish in indonesia
```{r}
## find all fish in Ind
indo_fish_ids <- index %>% filter(c_number == indo_code) %>% pull(taxonkey)

## select fish by taxonkey
fish_select_by_taxonkey <- taxa %>% filter(taxon_key %in% indo_fish_ids)

Fish <- read.csv("Indonesia_fisheries.csv") # combine code 200 & 203; repeat code from line 40 - 133 to get all fish in Indonesia
Yellowfin_tuna <- Fish %>% filter(fish_type == "Yellowfin_tuna")
Scombroids <- Fish %>% filter(fish_type == "Scombroids")
Skipjack_Tuna <- Fish %>% filter(fish_type == "Skipjack_tuna")

final_selection <- rbind(Yellowfin_tuna, Scombroids, Skipjack_Tuna)
write.csv(final_selection, file = "three_selected_Fish.csv")
```

# Perform some exploratory data analysis.
```{r EDA}
merged_effort <- read.csv("merged.csv")

merged_effort %>% 
ggplot(aes(x = average_bleaching, y = rate_norm)) +
  geom_point()

merged_effort %>% 
  mutate(bleached = ifelse(average_bleaching > 0, "Bleached", "Not bleached")) %>% 
  ggplot(aes(x = bleached, y = log(rate_norm))) +
  geom_boxplot()
```